---
output: github_document
always_allow_html: yes
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# familyR

The goal of familyR is to provide a fast way to get all parents, children and even the entire family tree of a node. It uses nothing but data.table and hence it is fast!

familyR consists of tree main functions:
1. get_familytree
2. get_parents
3. get_children

familyR is built so it is easy to visualize with visNetwork's hierarchical layout as both get_parents and get_children include the number of levels (or generations if you will), there are between the node you are looking and other nodes in the family.

## Installation

You can install familyR from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("emillykkejensen/familyR")
```

## Example

First load the package - note, that familyR relies heavily on data.table.

```{r}
library(familyR)
```

To use familyR you need a data.table of relations:

```{r}
my_network <- data.table::data.table(
  from = c("A", "A", "B", "B", "C", "C", "X", "Y"),
  to = c("B", "C", "D", "E", "E", "B", "Y", "Z")
)
```

Note that in order for it to work, you need to name the two columns "from" and "to" - you can include others if you like, as long as you have from and to!

Now let's try to get the entire family tree of node B.

```{r}
my_familytree <- get_familytree(my_network, "B")

print(my_familytree$nodes)

print(my_familytree$edges)
```

To show this with visNetwork all you need is:

```{r eval=FALSE}
library(visNetwork)

visNetwork(my_familytree$nodes, my_familytree$edges) %>%
  visEdges(arrows = "to")
```

If we on the other hand only want the parents of a node, use get_parents()

```{r}
my_parents <- get_parents(my_network, "D")

print(my_parents$nodes)

print(my_parents$edges)
```

As you can see the function not only returns the parents of a node, but also the grandparents, grand grandparents etc. and the function even includes the level (or number of generations) between the node you are looking for (in this case "D") and the other parent nodes in the family. 

Parents are given a negative value in level, whereas children are given a positive value.

To use this with visNetwork simply run:

```{r eval=FALSE}
visNetwork(my_parents$nodes, my_parents$edges) %>% 
  visEdges(arrows = "to") %>%
  visHierarchicalLayout()
```

To get all children of a node, use get_children()

```{r}
my_children <- get_children(my_network, "A")

print(my_children$nodes)

print(my_children$edges)
```

```{r eval=FALSE}
visNetwork(my_children$nodes, my_children$edges) %>% 
    visEdges(arrows = "to") %>%
    visHierarchicalLayout()
```

To show both parents and children combine the two functions with merge_family.

```{r}
my_parents <- get_parents(my_network, "C")

my_children <- get_children(my_network, "C")

my_family <- merge_family(my_parents, my_children)

print(my_family$nodes)

print(my_family$edges)
```

```{r eval=FALSE}
visNetwork(my_family$nodes, my_family$edges) %>% 
    visEdges(arrows = "to") %>%
    visHierarchicalLayout()
```

As you can see, some nodes are internally related (node B and E). If you would like to remove those inner relations use the remove_inner_relations() function.

```{r eval=FALSE}
my_new_family <- remove_inner_relations(my_family)

visNetwork(my_new_family$nodes, my_new_family$edges) %>% 
    visEdges(arrows = "to") %>%
    visHierarchicalLayout()
```


